---
title: "Simple document"
output: github_document
---

```{r}
library(tidyverse)
#incluing several packages,like reader, ggplot....
```


want to keep all things that in the same project, in this case, just create a new folder data within the same R folder and copy all import files into data folder----make sure all the data files that are useful for project in the same folder

let's import the `FAS_litters.cvs`csv using a relative path.

```{r}
litters_df=
  read_csv("data/FAS_litters.csv")
#readr package: sort by the library tidyverse, read_csv()functions;
#the first thing: let function where to find this file

#the paths show: when you just created a folder in your r project, this would show when you type its name
#using this relative path would not change even you transfer folder to other places

litters_df
#do to check if this works

litters_df = 
  janitor::clean_names(litters_df)
#in the cvs, there are spaces between names, but R cannot read them, we need to change names into snack_name
#this code make all names into snack_name format

#janior is a whole package in r, have bunch of functions in it

```


Import the same dataset using an absolute path.

```{r}
# problem: litters_df_abs=
#  read_csv("~/R Project/p8105_data_wrangle_1/data/FAS_litters.cvs")
#if you just use this absoluate path (when type and choose which folder that your data in), when you put the whole foler(include project and data into new folder,the path is changed, this would make mistake for this path)---get error(move folder)

# litters_df_abs


# litters_df_abs = 
#  janitor::clean_names(litters_df_abs)
```
```{r}
pups_df=
  read_csv("data/FAS_pups.csv")
pups_df = 
  janitor::clean_names(pups_df)

```


## look at data

```{r}
litters_df

head(litters_df)
#print the first six rows instead of first 10 rows

tail(litters_df)
#the bottom rows
```

you can use "view"
```{r}
view(litters_df)
#look at all the rows and columns
```


look at a data summary

```{r}
str(litters_df)
#loog at the structure of the file data fram until you sort of get used to how r works

skimr::skim(litters_df)
#gibe you some informations about data, with character and numeric variables---also give you what distribution looks like (histogram)
```

## options in `read_*`

```{r}
litters_df=
  read_csv(
    "data/FAS_litters.csv",
    skip = 10,col_names=FALSE)
#col_names=FALSE: does not show any name in columns, only x1,x2,x3..
```

look at NA values:
```{r}
litters_df=
  read_csv(
    "data/FAS_litters.csv",
    na = c("NA",19))

#the data  na = c("NA",19)) means that when everytime see the 19, miss this number


```


column types
```{r}
litters_df=
  read_csv(
    "data/FAS_litters.csv",
    col_types = 
      cols(
        Group = col_factor()
        #which colums = types 
      ))
#col_types is to change the types of column, like num>>>factor
litters_df

```


## other file types

import a xlsx file first

```{r}
library(readxl)
#reading xlsx file
```

```{r}
mlb_df=
  read_excel("data/mlb11.xlsx")
```

import a SAS file
```{r}
library(haven)
#read SAS
```

```{r}
pulse_df=
  read_sas("data/public_pulse_data.sas7bdat")
```

